---
title: "Inference interractions"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    code_folding: hide
    highlight: kate
bibliography: ref_interaction_inference.bib
---

<script src="https://hypothes.is/embed.js" async></script>

```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Note

Please, click on this link to get back to the [Home Page](index.html).

I will use two datasets : one from Stéphane Robin's training (Barents Fish).


```{r}
library(gllvm)
library(tidygraph)

library(corrplot)
library(gclus)
library(ggraph)
```

# Data

```{r}
load("BarentsFish.Rdata")

y <- as.data.frame(Data$count)

X <- as.data.frame(Data$covariates)
```

Looking at the abundance data.

```{r}
dim(y)

head(y)
```

Looking at the covariates.

```{r}
dim(X)

str(X)

summary(X)
```

Scaling environmental data.

```{r}
X <- as.data.frame(scale(X))
```

# Generalized Linear Latent Variable Model

## How does this package work?


A GLLVM model regress the mean abondance $u_{ij}$  of the specie $j$ at site $i$ as follow :

$$g(\mu_{ij}) = \eta_{ij} = \alpha_i + \beta_{0j} + x_i'\beta_j + u_i'\gamma_j$$
Where $\beta_j$ and $\gamma_j$ are vectors of species coefficient related to the covariates and to the factor loadings respectively.

Becaause there is some latent variables and factor loadings that there not observered (obvious), we estimate them with likelihood estimation.

How the algorithm work? First, `gllvm` fit a GLM model on each species. Then it computes *Randomized Quantile Residuals* also known as *Dunn-Smyth Residuals*. These residuals have the propertie to be exactly normaly distributed if model assumptions are correct [@Dunn_1996 ; @Niku_2019a]. That is why it allows the use of factor analysis using maximum-likelihood estimation to compute starting values for latent variables and factor loadings. Finaly, these values are passed into gradient based optimisation algorithms and give us back the factor loadings and latent variables values.

The residual covariance matrix stores informations about species co-occurence not explained by environmental variables computed as the product of loading matrix  the transposed loading matrix. 

$$\Sigma = \Gamma\Gamma'$$

Wich distribution is well suited for these data?

```{r}
fitp <- gllvm(y, family = poisson())

fitnb <- gllvm(y, family = "negative.binomial")
```

Look at the diagnostic plots.

```{r}
par(mfrow = c(3, 2))
plot(fitp, var.colors = 1)

par(mfrow = c(3, 2))
plot(fitnb, var.colors = 1)
```

Using *Poisson* distirubtion there is some overdispersion and there is not when using the *Negative Binomial* distribution. 

How many latent variable should we choose?

```{r}
AICc <- NULL
BIC <- NULL

for(i in 0:5){
  
  fiti <- gllvm(y, X, family = "negative.binomial", num.lv = i, sd.errors = FALSE, seed = 42)
  
  AICc[i+1] <- summary(fiti)$AICc
  BIC[i+1] <- summary(fiti)$BIC
  
  names(AICc)[i+1] = i
  names(BIC)[i+1] = i
}

print("AICc")
print(AICc)

print("BIC")
print(BIC)
```

The most parcimonious model is the one with only one latent variable.

Now we create a null model with one latent variable.

```{r}
fit_null <- gllvm(y, family = "negative.binomial", num.lv = 1, seed = 42)
```

And now the model

```{r}
fit <- gllvm(y, X, family = "negative.binomial", num.lv = 1, seed = 42)
```

How much of the variance is explained by the environmental variables?

```{r}
1 - getResidualCov(fit)$trace / getResidualCov(fit_null)$trace
```

Effect of environmental variable

```{r}
coefplot(fit, cex.ylab = 0.7, mfrow=c(1,1))
```

Most species of fish are affected by one or more environmental variable.

# Studying co-occurrence patterns

```{r}
cr <- getResidualCor(fit)
```

Plotting residual correlation

```{r}
corrplot(cr[order.single(cr), order.single(cr)], diag = FALSE, type = "lower",
method = "square", tl.cex = 0.5, tl.srt = 45, tl.col = "red")
```

Not so bad at all!!!

Now I will try to create a graph from this. I will keep only the 5% most correlated data (2.5% lower bound, 2.5% upper bound).

```{r}
cr_sub <- cr[lower.tri(cr)]

# Now I will keep only the most extreme correlation values

cr_sub <- cr_sub[cr_sub <= quantile(cr_sub, 0.025) | cr_sub > quantile(cr_sub, 0.975)]

cr_xtr <- cr
cr_xtr[!cr_xtr %in% cr_sub] <- 0

# Now I will remove rows and columns full of 0
cr_xtr <- cr_xtr[rowSums(cr_xtr) != 0, colSums(cr_xtr) != 0]

# Create the graph
graph <- as_tbl_graph(cr_xtr)
```

And now, a graph!

```{r}
gllvm_plot <- ggraph(graph, layout = "stress",) +
  geom_edge_link(aes(colour = weight), edge_width = 1.2) + 
  geom_node_point() + 
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_color_gradient2(name = "Residual\nCorrelation",
                             low  = "red",
                             high = "blue") +
  ggtitle("GLLVM method") +
  theme(panel.background = element_rect(fill = "white")) +
  theme_void()

gllvm_plot_viridis <- ggraph(graph, layout = "stress",) +
  geom_edge_link(aes(colour = weight), edge_width = 1.2) + 
  geom_node_point() + 
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_color_viridis(name = "Residual\nCorrelation") +
  ggtitle("GLLVM method") +
  theme(panel.background = element_rect(fill = "white")) +
  theme_void()

gllvm_plot
gllvm_plot_viridis
```

Ok, now I will use the Robin's model based on PLN models to reconstruct the same network. The code below come from one of his workshop. I do not understand quite well for the moment.

```{r}
library(PLNmodels)
library(EMtree)

# Create PLN model
plnNetAll <- PLNnetwork(as.matrix(y) ~ as.matrix(X))

plnNetAllBest <- getBestModel(plnNetAll)

plnNetAllG <- 1*(plnNetAllBest$model_par$Omega != 0) # We get interactions encoded by 0/1

# Compute Edge proba
treeAllWeight <- -nrow(y)/2*log(1-cov2cor(plnNetAllBest$model_par$Sigma)^2)
treeAllScore <- EdgeProba(treeAllWeight)

# Add convert binaries interaction into probability
graph_robin <- plnNetAllG * treeAllScore

graph_robin <- graph_robin[rowSums(graph_robin) != 0, colSums(graph_robin) != 0]

graph_robin <- as_tbl_graph(graph_robin)
```

```{r}
robin_plot <- ggraph(graph_robin, layout = "stress",) +
  geom_edge_link(aes(colour = weight), edge_width = 1.2) + 
  geom_node_point() + 
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_color_viridis(name = "Conditionnal\nprobabilities") +
  ggtitle("Robin's method") +
  theme(panel.background = element_rect(fill = "white")) +
  theme_void()

robin_plot
```

Comparing the two infered graph

```{r}
# Number of links
graph %>% activate(edges) %>% as_tibble() %>% nrow()
graph_robin %>% activate(edges) %>% as_tibble() %>% nrow()
```

Ok, so the `gllvm` method seem to infer more link compare to Robin's method (when I remove loops).

How many nodes are predicted by `gllvm` and Robin's method?

```{r}
## Extract the name of species
# GLLVM
node_name_gllvm <- graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  mutate(id = 1:nrow(.)) %>%
  mutate(id = as.character(.$id))

# Get a named vector for replacement
node_name_gllvm_vector <- node_name_gllvm$name %>%
  `names<-`(node_name_gllvm$id)

# Robin's Method
node_name_robin <- graph_robin %>%
  activate(nodes) %>%
  as_tibble() %>%
  mutate(id = 1:nrow(.)) %>%
  mutate(id = as.character(.$id))

# Get a named vector for replacement
node_name_robin_vector <- node_name_robin$name %>%
  `names<-`(node_name_robin$id)

# Compare species found by each metho
node_name_robin_vector[!node_name_robin_vector %in% node_name_gllvm_vector]

node_name_gllvm_vector[!node_name_gllvm_vector %in% node_name_robin_vector]

unname(node_name_robin_vector[node_name_robin_vector %in% node_name_gllvm_vector])
```
`gllvm` include 3 species in some interactions that there are not in Robin's method. On the other hand, Robin's method include 5 species in some interactions not predicted by `gllvm`. 9 species are both found by the two methods. It represents 75% shared species for `gllvm` and 64% shared species for Robin's method.

```{r}
## GLLVM
# Convert ID to species names
network_gllvm <- graph %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  mutate(from = as.character(from), to = as.character(to)) %>%
  mutate(from = stringr::str_replace_all(.$from, rev(node_name_gllvm_vector)), to = stringr::str_replace_all(.$to, rev(node_name_gllvm_vector))) # Use rev() otherwise, there 11 id is replace by "Re_hiRe_hi" and not by "Ly_va"

## Robin's Method
# Convert ID to species names
network_robin <- graph_robin %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  mutate(from = as.character(from), to = as.character(to)) %>%
  mutate(from = stringr::str_replace_all(.$from, rev(node_name_robin_vector)), to = stringr::str_replace_all(.$to, rev(node_name_robin_vector))) # Use rev() otherwise, there 11 id is replace by "Re_hiRe_hi" and not by "Ly_va"

network_gllvm[(network_gllvm$from %in% network_robin$from & network_gllvm$to %in% network_robin$to) | (network_gllvm$to %in% network_robin$from & network_gllvm$from %in% network_robin$to),]

network_gllvm_vector <- network_gllvm %>% 
  mutate(from_to = paste(.$from, .$to), to_from = paste(.$to, .$from)) %>%
  select(from_to, to_from) %>% 
  unlist(use.names = FALSE)

network_robin_vector <- network_robin %>%
  mutate(from_to = paste(.$from, .$to), to_from = paste(.$to, .$from)) %>%
  select(from_to, to_from) %>%
  unlist

sum(network_gllvm_vector %in% network_robin_vector)

sum(network_gllvm_vector %in% network_robin_vector) / (length(network_gllvm_vector) + length(network_robin_vector))
```

There are 16 interactions predicts by the two methods, it represents 12% of the whole interactions.

```{r}
network_common <- stringr::str_split_fixed(network_gllvm_vector[network_gllvm_vector %in% network_robin_vector], " ", n = 2)

graph_common <- as_tbl_graph(network_common)

common_plot <- ggraph(graph_common, layout = "stress",) +
  geom_edge_link(edge_width = 1.2) + 
  geom_node_point() + 
  geom_node_text(aes(label = name), repel = TRUE) +
  scale_edge_color_viridis() +
  ggtitle("Common edges between GLLVM and Robin's method") +
  theme(panel.background = element_rect(fill = "white")) +
  theme_void()

common_plot
```

Interestingly, it is not the edges with the highest conditional probabilities that are shared by these methods, but the edges with the highest residual correlation. Here the plots of the two methods.

# References
