---
title: "Inference interractions"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

<script src="https://hypothes.is/embed.js" async></script>

```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```


# List of packages 

List of packages to try :

- alien (Direct Matching Centrality, Indirect matching centrality, 4th corner, binary, Niche probabilistic, KNN as a minimal iEat)
- Hmsc (HMSC)
- boral (LVM)
- ecoCopula (Gaussian Copula Graphical Model)
- PLNmodels (Gaussian Graphical Model)
- Emtree (Tree-based inferrence + PLN)
- MInt (Poisson-multivariate normal hierachical model)
- MRFcov (Markov random fields)

# Packages & data

```{r}
# Packages

library(alien)
library(Hmsc)
library(boral)
library(ecoCopula)
library(PLNmodels)
library(EMtree)
library(MInt)
library(MRFcov)

# Data

load("Faune_cleaned_06-09-18.RData")
```

# Testing packages

## alien



## Hmsc



## boral


## ecoCopula


## Emtree



## MInt


## MRFcov