---
title: "Inference interractions"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

<script src="https://hypothes.is/embed.js" async></script>

```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```


# Note

Please, click on this link to get back to the [Home Page](index.html).

Here I will compare results of packages `HMSC`, `gllvm`, `ecoCopula` and `EMtree`.

# Packages & data

```{r}
# Packages

## Models

library(Hmsc)
library(ecoCopula)
library(PLNmodels)
library(EMtree)
library(gllvm)

## Graph
library(ggraph)
library(tidygraph)

## Plots
library(lattice)
library(corrplot)
library(gclus)
library(viridis)

# Data

data("spider")

sp_names <- colnames(spider$abund)

abund <- as.matrix(spider$abund)

env_sc <- data.frame(scale(spider$x))

## seed

set.seed(42)

## Number of chains for MCMC

nChains <- 3

## Number of cores for parallel computing

if(parallel::detectCores() > 2 & parallel::detectCores() > nChains){
  
  nParallel <- nChains
  
}else{
  
  nParallel <- 1
  
}
```

# HMSC

First, let's  try *HMSC*. We don't have informations about sampling design, let's assume that each site have been visited once.

```{r}
studyDesign <- data.frame(sample = as.factor(1:nrow(abund)))

rL <- HmscRandomLevel(units = studyDesign$sample)
```

Now, we will assume that there is no interactions between covariables.

```{r}
model_hmsc <- Hmsc(Y = abund, XFormula = ~soil.dry + bare.sand + fallen.leaves + moss + herb.layer + reflection, XData = env_sc, studyDesign = studyDesign, ranLevels = list(sample = rL))
```


Then we set up the *MCMC* chains.

```{r}
if(.Platform$OS.type != "windows"){
  thin <- 10
  samples <- 500
  transient <- 5*thin
  verbose <- 1
}else{
  thin <- 10
  samples <- 5500
  transient <- 50*thin
  verbose <- 1
}
```

And run the model.

```{r}
model_hmsc <- sampleMcmc(model_hmsc, thin = thin, samples = samples, transient = transient, nChains = nChains, verbose = 1, nParallel = nParallel)
```

Then we diagnostic the model.

```{r}
model_hmsc_post <- convertToCodaObject(model_hmsc)

par(mfrow=c(2,2))

hist(effectiveSize(model_hmsc_post$Beta), main = "ess(beta)")
hist(gelman.diag(model_hmsc_post$Beta, multivariate = FALSE)$psrf, main = "psrf(beta)")
hist(effectiveSize(model_hmsc_post$Omega[[1]]), main = "ess(omega)")
hist(gelman.diag(model_hmsc_post$Omega[[1]], multivariate = FALSE)$psrf, main = "psrf(omega)")
```

Let's see the correlation with the environnemtal variables.

```{r}
postBeta = getPostEstimate(model_hmsc, parName="Beta")
plotBeta(model_hmsc, post=postBeta, param="Support", supportLevel = 0.95)
```

There is no correlation with the species and the bare sediment and reflexion. 

```{r}
OmegaCor <- computeAssociations(model_hmsc)
supportLevel <- 0.95

toPlot <- ((OmegaCor[[1]]$support > supportLevel) + (OmegaCor[[1]]$support < (1-supportLevel)) > 0) * OmegaCor[[1]]$mean

corrplot(toPlot, method = "color", col = colorRampPalette(c("blue","white","red"))(200),
title = paste("random effect level:", model_hmsc$rLNames[1]), mar=c(0,0,1,0))
```

So there is no inter-specific interactions.

# gllvm

First, let's choose the number of latent variables.

```{r}
criterias <- NULL

for(i in 0:5){
  
  fiti <- gllvm(abund, env_sc, family = "negative.binomial", num.lv = i, formula = abund ~ soil.dry + bare.sand + fallen.leaves + moss + herb.layer + reflection)
  
  criterias[i+1] <- summary(fiti)$AICc
  
  names(criterias)[i+1]= i
}

print(criterias)
```

Based on $AIC$, We will take 1 latent variable. Now we are fitting the model.

```{r}
model_gllvm <- gllvm(abund, env_sc, family = "negative.binomial", num.lv = 1, formula = abund ~ soil.dry + bare.sand + fallen.leaves + moss + herb.layer + reflection)
```

Let's see some diagnostic plots.

```{r}
par(mfrow = c(3, 2), mar = c(4, 4, 2, 1))
plot(model_gllvm, var.colors = 1)
```



Effect of environnemtal variables.

```{r}
coefplot(model_gllvm, cex.ylab = 0.7, mfrow = c(1,1))
```

Species interactions

```{r}
cr0 <- getResidualCor(model_gllvm)

corrplot(cr0[order.single(cr0), order.single(cr0)], diag = FALSE, type = "lower", method = "square", tl.cex = 0.8, tl.srt = 45, tl.col = "red")
```

Besides one specie (Arctperi), all have negative residual correlation.

# ecoCopula

```{r}
GCGM_mod <- manyglm(mvabund(abund) ~ as.matrix(env_sc))
GCGM_graph <- cgr(GCGM_mod)

plot(GCGM_graph, pad=1)
```

It's making a small graph, some species do not seems to interact.


# EMTree

We'll start with the same model

```{r}
EMT_model <- PLN(abund ~ as.matrix(env_sc))

p <- ncol(abund) # Number of nodes

EMT_model_res <- EMtree(EMT_model, maxIter = 10, plot = TRUE)

str(EMT_model_res)
```


```{r}
edges_prob <- EMT_model_res$edges_prob

edges_prob[edges_prob < 2/p] <- 0 # Remove links too improbable

EMtree::draw_network(edges_prob, nodes_label = sp_names, pal="dodgerblue3",layout="nicely", curv=0.1)
```